Structural analysis

### 2024-07-12 workflow/notebooks/draft.Rmd
fixed apro

pident_p <- ggplot(df, aes(pident_Blast, pident_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  scale_fill_gradientn(colors=wespal) +
  # scale_fill_viridis_c(option = "C") +
  coord_fixed(expand = 0) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(legend.position = "none", panel.grid = element_blank())
pident_p <- add_marginal(df, pident_p, "pident_Blast", "pident_Foldseek", "singleton", palette_singleton_p)

qcov_p <- ggplot(df, aes(qcov_Blast, qcov_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  geom_hline(yintercept = min_covs, linetype=2) + 
  geom_vline(xintercept = min_covs, linetype=2) + 
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_fill_gradientn(colors=wespal) +
  coord_fixed(expand = 0) +
  theme(legend.position = "none", panel.grid = element_blank()) 
qcov_p <- add_marginal(df, qcov_p, "qcov_Blast", "qcov_Foldseek", "singleton", palette_singleton_p)

lddt_p <- df %>% 
  select(lddt, singleton) %>% 
  ggplot(aes(color=singleton, lddt)) + 
  geom_density() + 
  # scale_y_continuous(expand = c(0.01,0)) +
  # scale_x_continuous(limits = c(params$low_confidence, 100),
  #                    expand = c(0,0)) +
  scale_color_manual(values = palette_singleton_p)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

alntmscore_p <- df %>% 
  select(alntmscore, singleton) %>% 
  ggplot(aes(color=singleton, alntmscore)) + 
  geom_density() + 
  # coord_cartesian(expand = 0) +
  scale_color_manual(values = palette_singleton_p)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

plddt_p <- df %>% 
  select(target, singleton) %>% 
  mutate(target=gsub("AF-|-F1", "", target)) %>% 
  left_join(lddt_df, by="target") %>% 
  ggplot(aes(color=singleton, mean_lddt)) + 
  geom_density(bw=1) + 
  # scale_y_continuous(expand = c(0.01,0)) +
  scale_x_continuous(limits = c(params$low_confidence, 100)) +
  scale_color_manual(values = palette_singleton_p)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

CATH_plot <- df_multi %>% 
    pivot_longer(contains("Jaccard"), values_to = "Jaccard") %>% 
    mutate(name=gsub("Jaccard_", "", name),
           name=gsub("3d", "CATH", name)) %>% 
    ggplot(aes(Jaccard, fill=singleton)) +
    geom_histogram(binwidth = 0.1, position = "dodge") +
    facet_grid(name~.) +
    scale_x_continuous(breaks = seq(0,1,0.1)) +
    scale_y_continuous(expand = expansion(0,0)) + 
    scale_fill_manual(values=palette_singleton_p) + 
  theme(legend.position = "bottom")

@@

fig2_first_row <- (pident_p | eval_p | qcov_p)
fig2_second_row <- ((lddt_p / alntmscore_p / plddt_p) | CATH_plot) + plot_layout(widths = c(1,2))
fig2 <-  (fig2_first_row / fig2_second_row)

ggsave("draft/figures/fig2.pdf", fig2, width = 9, height = 6)
```

@@

```{r fig3}
my_comparisons <- list( c("LG", "3Di"), c("3Di", "GTR"), c("LG", "GTR") )
plot_bs <- df_bs %>% 
  # group_by(gene, target, model) %>% 
  # summarise(mn_bs=mean(support)) %>% 
  ggplot(aes(model, support, fill=target)) +
  # geom_violin(position = position_dodge(width = 1)) +
  geom_boxplot() +
  # ggdist::stat_pointinterval(position = position_dodge(width = 1), color="black") +
  scale_fill_manual(values = palettes_method) +
  theme_classic() +
  labs(y="Boostrap", x="") +
  theme(legend.position = "none") 
  # ggpubr::stat_compare_means(aes(group = model),
  #                            comparisons = my_comparisons,
  #                            label = "p.format")

rf_plot <- rf_df %>% 
  ggplot(aes(RF)) +
  geom_tile(aes(x=0.5, y=0.5, fill=value), data = rf_text) +
  geom_text(aes(x=0.5, y=.5, label=round(value, 2)), color="white", data = rf_text) +
  geom_rug(linewidth=.1, data= . %>% filter(as.integer(model.x)>as.integer(model.y))) +
  geom_density(aes(y = after_stat(scaled)), data= . %>% filter(as.integer(model.x)>as.integer(model.y))) + 
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Median RF") + 
  scale_x_continuous(breaks = c(0,.25,.5, .75,1), labels = function(x) sprintf("%.2g", x)) +
  coord_fixed(xlim = c(0,1), expand=0) +
  # facet_grid(method~method2, switch = "both") +
  facet_grid(model.x~model.y) +
  # ggh4x::facet_nested(targets+model~targets2+model2, nest_line = element_line(linetype = 1)) +
  labs(x="RF Distance") +
  theme_classic() +
  theme(legend.position = "bottom", 
        panel.background = element_rect(fill='transparent'),
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip.text = element_text(size=6),
        axis.text.x = element_text(size=6),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())

@@

plot_DL <- scores %>% 
  filter(model!="FTPY") %>% 
  ggplot(aes(model, n_events, fill=targets)) +
  geom_boxplot() +
  labs(x="", y="(D+L)/# Tips") +
  # ggthemes::scale_fill_solarized() +
  scale_fill_manual(values = palettes_method) +
  theme(legend.position = "none")

var_r2t <- scores %>% 
  filter(model!="FTPY", targets=="common") %>% 
  ggplot(aes(color=model, variance_r2t)) + 
  geom_density(size=1, alpha=.5) +
  # geom_histogram(position = "dodge") + 
  coord_cartesian(expand = 0) +
  scale_x_continuous(limits = c(0,.2)) + 
  # facet_grid(~targets) + 
  labs(x="Variance Root-to-tip") +
  scale_color_manual(values = palettes_model) +
  theme(legend.position = "bottom")

fig3 <- rf_plot | ((plot_bs | plot_DL) / var_r2t)
ggsave("draft/figures/fig3.pdf", fig3, width = 11, height = 6)
```

@@

sptree_plot <- ggtree(labeled_sptree) + 
  geom_tiplab() + 
  geom_nodelab(color="firebrick", 
               nudge_x = -.1, hjust = 1, 
               nudge_y = .2, size=2)+
  scale_x_continuous(expand = c(.2,.2))

fig_s1 <- (plot_dist + sptree_plot + pp1_plot) + plot_layout(widths = c(1.4, 1, 1.6))

ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 6)
```

>>>>

pident_p <- ggplot(df, aes(pident_Blast, pident_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  scale_fill_gradientn(colors=wespal) +
  # scale_fill_viridis_c(option = "C") +
  coord_fixed(expand = 0) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="Identity Blast",y="Identity Foldseek") +
  theme(legend.position = "none", panel.grid = element_blank())
pident_p <- add_marginal(df, pident_p, "pident_Blast", "pident_Foldseek", "singleton", palette_singleton_p)

qcov_p <- ggplot(df, aes(qcov_Blast, qcov_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  geom_hline(yintercept = min_covs, linetype=2) + 
  geom_vline(xintercept = min_covs, linetype=2) + 
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="Query cov. Blast",y="Query cov. Foldseek") +
  scale_fill_gradientn(colors=wespal) +
  coord_fixed(expand = 0) +
  theme(legend.position = "none", panel.grid = element_blank()) 
qcov_p <- add_marginal(df, qcov_p, "qcov_Blast", "qcov_Foldseek", "singleton", palette_singleton_p)

lddt_p <- df %>% 
  select(lddt, singleton) %>% 
  ggplot(aes(color=singleton, lddt)) + 
  geom_density() + 
  # scale_y_continuous(expand = c(0.01,0)) +
  # scale_x_continuous(limits = c(params$low_confidence, 100),
  #                    expand = c(0,0)) +
  scale_color_manual(values = palette_singleton_p) + 
  labs(x="LDDT") + 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

alntmscore_p <- df %>% 
  select(alntmscore, singleton) %>% 
  ggplot(aes(color=singleton, alntmscore)) + 
  geom_density() + 
  # coord_cartesian(expand = 0) +
  labs(x="TM-Score") + 
  scale_color_manual(values = palette_singleton_p)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

plddt_p <- df %>% 
  select(target, singleton) %>% 
  mutate(target=gsub("AF-|-F1", "", target)) %>% 
  left_join(lddt_df, by="target") %>% 
  ggplot(aes(color=singleton, mean_lddt)) + 
  geom_density(bw=1) + 
  # scale_y_continuous(expand = c(0.01,0)) +
  scale_x_continuous(limits = c(params$low_confidence, 100)) +
  scale_color_manual(values = palette_singleton_p) +   
  labs(x="Mean target pLDDT") + 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

CATH_plot <- df_multi %>% 
    pivot_longer(contains("Jaccard"), values_to = "Jaccard") %>% 
    mutate(name=gsub("Jaccard_", "", name),
           name=gsub("3d", "CATH", name),
           Jaccard=cut_width(Jaccard, .1, boundary = 0)) %>% 
  filter(!is.na(Jaccard)) %>% 
  group_by(name, singleton, Jaccard) %>% 
  summarise(n=n()) %>% 
  mutate(freq=n/sum(n)) %>%
  ggplot(aes(Jaccard, freq, fill=singleton)) + 
  geom_bar(stat = "identity",position = "dodge") + 
  facet_grid(name~.) +
  labs(y="% in group") +
  scale_y_continuous(expand = expansion(0.01,0)) + 
  scale_fill_manual(values=palette_singleton_p) +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill=NA, color="black"),
        axis.line = element_blank(),
        axis.text.x = element_text(size=8))

@@

fig2_first_row <- (pident_p | eval_p | qcov_p)
fig2_second_row <- ((lddt_p / alntmscore_p / plddt_p) | CATH_plot) + plot_layout(widths = c(1,2))
fig2 <-  (fig2_first_row / fig2_second_row) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

ggsave("draft/figures/fig2.pdf", fig2, width = 10.5, height = 8)
```

@@

```{r fig3}
my_comparisons <- list( c("LG", "3Di"), c("3Di", "GTR"), c("LG", "GTR") )
plot_bs <- df_bs %>% 
  mutate(targets=target) %>% 
  # group_by(gene, target, model) %>% 
  # summarise(mn_bs=mean(support)) %>% 
  ggplot(aes(model, support, fill=targets)) +
  # geom_violin(position = position_dodge(width = 1)) +
  geom_boxplot(outlier.size = .1) +
  # ggdist::stat_pointinterval(position = position_dodge(width = 1), color="black") +
  scale_fill_manual(values = palettes_method) +
  theme_classic() +
  labs(y="Boostrap", x="") 
  # theme(legend.position = "bottom") 
  # ggpubr::stat_compare_means(aes(group = model),
  #                            comparisons = my_comparisons,
  #                            label = "p.format")

rf_plot <- rf_df %>% 
  ggplot(aes(RF)) +
  geom_tile(aes(x=0.5, y=0.5, fill=value), data = rf_text) +
  geom_text(aes(x=0.5, y=.5, label=round(value, 2)), color="white", data = rf_text) +
  geom_rug(linewidth=.1, data= . %>% filter(as.integer(model.x)>as.integer(model.y))) +
  geom_density(aes(y = after_stat(scaled)), data= . %>% filter(as.integer(model.x)>as.integer(model.y))) + 
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Median RF") + 
  scale_x_continuous(breaks = c(0,.25,.5, .75,1), labels = function(x) sprintf("%.2g", x)) +
  coord_cartesian(xlim = c(0,1), expand=0) +
  # facet_grid(method~method2, switch = "both") +
  facet_grid(model.x~model.y) +
  # ggh4x::facet_nested(targets+model~targets2+model2, nest_line = element_line(linetype = 1)) +
  labs(x="RF Distance") +
  theme_classic() +
  theme(legend.position = "bottom", 
        panel.background = element_rect(fill='transparent'),
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip.text = element_text(size=6),
        axis.text.x = element_text(size=6),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-20,0,0,0))

@@

plot_DL <- scores %>% 
  filter(model!="FTPY") %>% 
  ggplot(aes(model, n_events, fill=targets)) +
  geom_boxplot(outlier.size = .1) +
  labs(x="", y="(D+L)/# Tips") +
  # ggthemes::scale_fill_solarized() +
  scale_fill_manual(values = palettes_method)
  # theme(legend.position = "bottom")

var_r2t <- scores %>% 
  filter(model!="FTPY", targets=="common") %>% 
  ggplot(aes(color=model, variance_r2t)) + 
  geom_density(size=1, alpha=.5) +
  # geom_histogram(position = "dodge") + 
  # coord_cartesian(expand = 0) +
  scale_x_continuous(limits = c(0,.2)) + 
  # facet_grid(~targets) + 
  labs(x="Variance Root-to-tip") +
  scale_color_manual(values = palettes_model) +
  theme(legend.position = 'bottom',
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,0,0))

secondcolumn <- (plot_DL + plot_bs + 
    plot_layout(guides = "collect", widths = c(5,3)) &
    theme(legend.position = 'bottom',
          legend.margin=margin(0,0,0,0),
          legend.box.margin=margin(-20,0,0,0))) / var_r2t

fig3 <- (rf_plot | secondcolumn) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
ggsave("draft/figures/fig3.pdf", fig3, width = 11, height = 6)
```

@@

nodes_df %>% 
  ggplot(aes(label_int, f1+f2+f3, fill=model)) + 
  geom_bar(stat="identity", position = "dodge") +
  facet_grid(targets~.) +
  scale_fill_manual(values=palettes_model)

sptree_plot <- ggtree(labeled_sptree) + 
  geom_tiplab() + 
  geom_nodelab(color="firebrick", 
               nudge_x = -.15, hjust = 1, 
               nudge_y = .3, size=2)+
  scale_x_continuous(expand = c(.2,.2))

plot_ntips <- scores %>% 
  filter(model!="FTPY") %>% 
  ggplot(aes(n_tips, dups+losses, color=targets)) + 
  geom_point(size=.2) +
  geom_smooth(method = "lm") +
  facet_grid(~model) + 
  scale_color_manual(values = palettes_method) +
  theme_classic() + 
  theme(panel.border = element_rect(fill=NA),
        legend.position = "bottom")

fig_s1 <- (plot_dist + sptree_plot + pp1_plot) +
  plot_layout(widths = c(1.4, 1, 1.6)) + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 6)

# fig_s1 <- (plot_ntips / ((plot_dist + sptree_plot + pp1_plot) +
#                            plot_layout(widths = c(1.4, 1, 1.6)))) + 
#   plot_layout(heights = c(.75,1)) 
# ggsave("draft/figures/figS1.pdf", fig_s1, width = 14, height = 8)
```

## Answers to Edu's comments

```{r}
df %>% 
  filter(query %in% seeds) %>% 
  count(query, singleton) %>% 
  ggplot(aes(n, singleton, fill=singleton)) +
  geom_boxplot()

blast_filtered_eval <- filter(blast, 
                         query %in% paste0("AF-", aln_seeds, "-F1"),
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

fs_filtered_eval <- filter(fs, 
                      query %in% paste0("AF-", aln_seeds, "-F1"),
                      evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

df_filtered_eval <- full_join(blast_filtered_eval, fs_filtered_eval,
                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_Foldseek = 100*pident_Foldseek, 
         qcov_Foldseek = 100*qcov_Foldseek,
         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
                               is.na(evalue_Blast) ~ "only Foldseek", 
                               .default = "Common"))

select(df_filtered_eval, query, target, singleton) %>% 
  mutate(method="only-eval") %>% 
  rbind(select(df_filtered, query, target, singleton) %>% 
          mutate(method="qcov-tcov")) %>% 
  group_by(method, singleton) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

res %>% 
  group_by(singleton, model) %>% 
  summarise(median(nn_dist/max_dist))
```
